---
title: "Penguins_ANOVA"
author: "Alex Brown"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

```{r, include=FALSE}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")


```

Our question:

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, and Adelie)?

2. Do some exploring 

New Functions: with + table for easy counting over multiple factors.

Update species name using case_when

```{r}
counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>%
  mutate(common_name =
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
           )) %>% 
  select(-species) %>% 
  filter(sex == "FEMALE")
```

- Exploratory Graphs
- Summary Statistics
- Levene's tests for equal variances
- one-way ANOVA
- post-hoc testing

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~ common_name)

hists

qqs <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() + 
  facet_wrap(~ common_name)

qqs


# Levene's test for equal variances (> 2 groups...if you have two groups, use the F-test for equal variances)

# we'll use the function leveneTest from the 'car' package

# HO: variances are equal
# HA; variances are not equal

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)

penguin_levene

# We retain the null hypothesis of equal variances (p = 0.06)

variances <- penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    variance = var(flipper_length)
  )

variances

```

3. One-way ANOVA

Single factor: common_name (species)
Number of levels: 3 (chinstrap, Gentoo, Adelie)
Random variable: flipper_lenght (mm)

Is there a significant difference in mean flipper length (mm) for female Chinstrap, Gentoo, and Adelie penguins? 

```{r}
# HO: mean flipper lengths across all species are equal
# HA: At least two means differ significantly 

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

summary(penguin_aov)

# at least two samples were taken from populations with different means. Which ones are different? All three are different from each other? Or something else?


# post-hoc testing using Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)

penguin_ph
```

Mean flipper lengths for Adelie, Gentoo, and Chinstrap penguins are significantly differerent and are all significantly different from each other. (F(df) = statistic, p<0.001, $\alpha$ = .05 with post-hoc Tukey's HSD, $\alpha = 0.05$.
